---
title: "R"
author: Pai Jing-Yun
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to R
### Basic features in R
We tell R to make the assignment using an arrow “<-”. R also supports using “=” in most circumstances. However, “<-” is recommended as it makes clear that we are requesting an assignment, rather than stating a relation.
```{r}
a <- 3 ^ 5
b = 3 ^ 5
a
b
```
If you don’t assign a value to a name of variable, then you will get the error message.

### Special values in R
Missing value, Not a Number, Infinity, π, and 10−3
```{r}
NA
NaN ### 0/0
Inf ### 2/0
-Inf ### -2/0
pi
1e-3
```
### Name an objuect
+ Avoid the following words:
  - break; else; FALSE; F; TRUE; F; for; function; if; Inf; NA; NaN; next; repeat; return; while;
+ Name a meaningful object by “.” or "_".
  - score1, score2
  - score.John, score.Jack

### More on calculation
Absolute value: |x|
```{r}
abs(-2)
```
Exponential & logarithm: exp(x), log(x)
```{r}
exp(10); log(2); log2(2); log(10); log10(10)
log(10, base = 10); log10(10);log(10, base = exp(1)); log(10)
```
“ˆ” or "**" means the degree of the power
```{r}
2 ** 3; 2^3;2^0.5; sqrt(2)
```
Trigonometric function:
sin(), cos(), tan(), asin(), acos(), atan()

### R is not just a calculator
There are a lot of functions you can use. Let’s see some related to Statistics.
```{r}
x <- 1:25
mean(x)
var(x)
median(x)
quantile(x, c(0.1, 0.5, 0.7))
```
We use c() to be a collection of needed values
```{r}
x <- c(1, 3, 5, 7, 9, 11)
y <- c(2, 4, 6, 8, 10, 12)
x
y
```
Easy way to get the summary of a set
```{r}
summary(x)
```
### Figures of a set
```{r}
x <- c(1:25, NA)
par(mfrow = c(1, 2))
boxplot(x)
hist(x)
```

# Introduction to vector
```{r}
vt1 <- c(2:4)
vt2 <- c(3:6)
vt2 <- c(vt2, vt1)
vt2
vt2 <- c(vt2, vt1)
vt2
```

### Some common functions to vectors
```{r}
length(vt1)
length(vt2)
```
Extracting elements from vectors, use [ ] to select the locations in the vector. Note that the count is different from C and python.
```{r}
vt1[3]
vt2[c(3,5,7)]
```
We can select all but the some elements of the vector as follows:

```{r}
vt1
vt1[-2]
vt2[-c(3:6)]
x <- c(3, 6, 9)
x * 3
x - 3
x ^ 3
y1 <- c(3, 2); y2 <- c(3, 2, 4);
x^y1
x^y2
```
When the vectors are different lengths, the shorter one is extended by recycling
```{r}
c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 7, 7, 8, 8, 9, 9, 10, 10) %% 2:3
```
Recycling: rep(x, ntimes)
```{r}
c(1, 1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 7, 7, 8, 8, 9, 9, 10, 10) %% rep(2:3, 10)
```
Use “seq()” and “rep()” to construct a vector with similar patterns.
```{r}
rep(3, 12)
rep(seq(2, 10, 2), 2)
rep(c(1, 4), c(3, 2))
rep(seq(2, 10, 2), rep(3, 5))
Xvector <- rep(NA, 6)
Xvector[3] <- 12
Xvector[c(4:6)] <- rep(1, 3)
Xvector
```
How to delete NA?
```{r}
Xvector[-(1:2)]
Xvector <- Xvector[-(1:2)]
Xvector
```
+ When to use “c( )” and when to use a vector with known length?
  - c( ): Easy to add a value into the vector; need not know the length.
  - rep(NA, n): With known length and the position of elements.

Use “unique( )” to return the elements that appear at least one time.
Use “table( )” to count the number of each element from the vector.
Let b = {3, 3, 3, 3, 5, 5, 1, 2, 10, 10, 8}
```{r}
b <- c(3, 3, 3, 3, 5, 5, 1, 2, 10, 10, 8)
unique(b)
table(b)
```
Set Operations
```{r}
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
union(B, C)
intersect(B, C)
setdiff(B, C)
setequal(B, C)
```
From small to large values, use “sort( )” or “order( )” to sort a vector or get the position of sorting of a vector.
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
sort(a)
sort(a, decreasing = TRUE)
```
Get the position (index) of sorting of a vector
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
inc.order <- order(a)
dec.order <- order(a, decreasing = TRUE)
inc.order
a[inc.order]
dec.order
a[dec.order]
```

### Change values in a vector
Let a = {3, 4, 2, 7, 10, 23, 8, 9, 11}. Change the 5th element to -3 and the 8th element to -1.
```{r}
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
a[5] <- -3
a[8] <- -1
a
vt1
vt1[-2]
vt2[-c(3:6)]
```
### Character vectors
```{r}
"Hello World!!!"
print("Hello R World!!!")
TEXT1 <- "Let's go to a game."
strsplit(TEXT1, " ")
TEXT2 <- "Hurry! We are late!"
strsplit(TEXT2, "!")
toupper(TEXT1)
tolower(TEXT2)
colors <- c("red", "yellow", "BLUE")
more.colors <- c(colors, "green", "cyan")
more.colors
z <- c("red", "green", 12)
z
```

### Basic operations for character vectors
-To take substrings, use “substr( )”
```{r}
colors <- c("red", "yellow", "blue")
substr(colors, 1, 3)
```
To search a string from a vector, use “grep()”, which is “Global Regular Expression Print”. It returns the locations of the desired patterns
```{r}
grep("y", more.colors)
grep("e", more.colors)
y.in.V <- grep("y", more.colors)
e.in.V <- grep("e", more.colors)
more.colors[y.in.V]
more.colors[e.in.V]
E.in.V <- grep("E", more.colors)
more.colors[E.in.V]
```
-To building up strings, use “paste()”.
```{r}
paste(colors, "line") ## with space
paste(colors, "line", sep = "" ) ## no space
## use ";" to seperate strings
paste("I like", colors, collapse = "; ")
```
factors
```{r}
colors <- c("red", "yellow", "blue")
f.c <- factor(colors)
f.c
levels(f.c)
as.integer(f.c)
```
### logical vector
TRUE and FALSE are reserved words denoting logical constants in the R, whereas T and F are global variables whose initial values set to these.
Binary operators which allow the comparison of values in atomic vectors (relational operators).
```{r}
c(3 < 5, 3 > 5)
c(3 <= 5, 3 >= 5)
3 == 5
3 != 5
#Boolean operations
A <- c(T, T, F, T, F); B <- c(3, 1, 5, 4, 7)
A
B[A]
!A
B[!A]
#How many TRUE (or FALSE) in the vector?
A <- c(T, T, F, T, F)
sum(A)
sum(!A)
#Compare the same location of elements individually
B <- c(3, 1, 5, 4, 7); C <- c(23, 1, 44, 5, 3)
B == C
D <- c(2, 5 ,6, 10)
B == D
```
### “is.” series
```{r}
is.null(c())
c(is.na(NA), is.na(2))
x1 <- 1:5
x2 <- c(1:3, 2.2)
x3 <- colors
c(is.integer(x1), is.numeric(x1), is.double(x1))
c(is.integer(x2), is.numeric(x2), is.double(x2))
c(is.character(x3), is.numeric(x3))
#Use “class( )” to identify the type of a vector
a <- c(3, 4, 2, 7, 10, 23, 8, 9, 11)
b <- c("red", "yellow", "blue")
c <- c(TRUE, FALSE)
d <- f.c
c(class(a), class(b), class(c), class(d))
c(class(2), class(2.0), class(as.integer(2)))
```
# Matrix
### Construct a matrix
A matrix is a generalization containing multiple vectors.
```{r}
a <- 1:5; b <- 6:10;
c <- 11:15; d <- 16:20;
M1 <- rbind(a, b, c, d)
M1
M2 <- cbind(a, b, c, d)
M2
dim(M1)
dim(M2)
X <- 1:20
matrix(X, nrow = 4, ncol = 5)
matrix(X, nrow = 5, ncol = 4)
matrix(X, nrow = 4, ncol = 5, byrow = TRUE)
matrix(X, nrow = 5, ncol = 4, byrow = TRUE)
```
###　Features for a matrix
```{r}
M1
dim(M1)
colnames(M1); rownames(M1);
colnames(M1) <- c("V1", "V2", "V3", "V4", "V5")
dimnames(M1)
dimnames(M2)
```
We still use “[ ]” to extract elements from a matrix, but using a “,” to identify the locations of row and column seperately.
```{r}
M1[2, 4];M1[c(2, 4), 3]; M1[c(2, 4), c(3, 5)];M1[, 4];
M1[c(2, 4), ]; M1[, c(3, 5)]
#Extract elements by names
M1[c("b", "c"), ]
#The names of columns are V1, V2, V3, V4, and V5. Take the rows of b and c and the columns of V2 and V5.
```

# Control Flow
### For() loop
+ When to use?
  - See patterns of behaviour, not just a single instance.
  - The for() statement allows one to specify that a certain operation should be repeated a fixed number of times.
+ Syntax
  - for (index in vector) commands,
+ where
  - index: the index is changed each time
  - vector: it includes all possible cases
  - commands: actions

Construct a vector as
___**Xn+1 = 2Xn + 3**___
where X0 = 3, n = 1, 2, 3, 4, 5, 6, 7.
```{r}
vec.X.v1 <- c()
Xn <- 3
for(n in 1:7){
Xn <- 2*Xn + 3
vec.X.v1 <- c(vec.X.v1, Xn)
}
vec.X.v1
```
Evaluate n!
**n! = n ∗ (n − 1) ∗ (n − 2) ∗ · · · ∗ 2 ∗ 1.**
Try n = 10.
```{r}
n <- 10
ans <- 1
for(j in n:1){
ans <- ans * j
}
ans
factorial(n)
```
### Random number generator
Random samples from [0, 1] by the linear congruntial generator.
___**xn = (axn−1 + c) mod m.**___
Let x0 = 3, a = 5, c = 7, and m = 501. Find xi , where i = 1, . . . , n.  
What are the values after all in Xn devided by m?
```{r}
x0 <- 3; a <- 5; c <- 7; m <- 501
X <- rep(NA, 10)
for(j in 1:10){
x0 <- (a*x0+c) %% m
X[j] <- x0
}
X
round(X/m, 2)

x0 <- 3; a <- 5; c <- 7; m <- 501
X <- rep(NA, 11)
X[1] <- x0
for(j in 1:10){
X[j + 1] <- (a*X[j] + c) %% m
}
X[2:11]
round(X/m, 2)
```

### The if() statement
+ Syntax
+ if (condition) {commands when TRUE}
+ if (condition) {commands when TRUE} else {commands when FALSE}
  - This statement causes a set of commands to be invoked if condition evaluates to TRUE.
  - The else part is optional, and provides an alternative set of commands which are to be invoked in case the logical variable is FALSE.
Try to find the locations of NA
```{r}
C <- c(3, 4, NA, 12, 0.2, NA, -2)
for(i in 1:7){
if(is.na(C[i])) print(i)
}
na.loc <- c() 
for(i in 1:7){
if(is.na(C[i])) na.loc <- c(na.loc, i)
}
na.loc
```
### The while() loop
The pattern of repetition isn’t known in advance and we need to do some calculations and keep going as long as a condition holds. The while()statement accomplishes this.  
**Syntax**  
while (condition) statements  
If the condition evaluates to FALSE, nothing more is done. If it evaluates to TRUE the statements are executed, condition is evaluated again, and the process is repeated.  
while(): Need the initial, condition of stopping, and iteration or replacement  
```{r}
j <- 1
while(j <= 8){
print(j)
j <- j + 1
}
```
### Break, repeat and next
+ break
  - if (condition) break 
  The break statement causes the loop to terminate immediately. break statements can also be used in for() and while() loops.
+ repeat
  - Use “repeat” to repeat until we execute a break statement.
  repeat { statements }
  This causes the statements to be repeated endlessly.
+ next:
  - The next statement causes control to return immediately to the top of the loop; it can also be used in any loop.
```{r}
#List all Fibonacci numbers less than 100. Use for() and if(condition) break
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
for(j in 1:1000){
if(Fib2 < 100){
Fibonacci <- c(Fibonacci, Fib2)
Fib1 <- Fibonacci[length(Fibonacci) - 1]
Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}else{
break
}
}
Fibonacci
#List all Fibonacci numbers less than 100. Use repeat{ } and break.
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
repeat
{
if(Fib2 < 100){
Fibonacci <- c(Fibonacci, Fib2)
Fib1 <- Fibonacci[length(Fibonacci) - 1]
Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
}else{
break
}
}
Fibonacci
#List all Fibonacci numbers less than 100. Use repeat{ } and next.
Fib1 <- 1
Fib2 <- 1
Fibonacci <- c(Fib1)
for(j in 1:100){
if(Fib2 >= 100) next
Fibonacci <- c(Fibonacci, Fib2)
Fib1 <- Fibonacci[length(Fibonacci) - 1]
Fib2 <- Fib1 + Fibonacci[length(Fibonacci)]
#print(j)
}
Fibonacci
```
# Functions
1. Creat a new script for your function or codes.
2. Save the script named as “XXX.R” and use source( ) to run the code.
3. Remember to use print( ) or cat( ) so that your final answer will be
shown.
```{r}
x <- c(0.8933, 0.2344, 0.0764, 0.7736)
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
print(x.show)
cat(x.show)
#Find all perfect numbers smaller than 100. Show the results by print( ).
k <- 1
while(k < 100){
fac <- c()
for(j in 1:k){
if((k %% j) == 0) fac <- c(fac, j)
}
if(k == sum(fac[-length(fac)])){
out <- paste(paste(k, "=", sep=" "),
paste(fac[-length(fac)], collapse = " + "),
sep = " ")
print(out)
}
k <- k+1
}
```
### function()
```{r}
f <- function(x){
3*x + 2
}
f(2); f(3); f(4)
make.percent <- function(x){
x.percent <- round(100*x, digits = 1)
x.show <- paste(x.percent, "%", sep = "")
cat(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736))
#You must tell what will be return when you use the function by return( ) or invisible( ). If you don’t put return( ) at the end of the function, it will treat the last line as the result automatically.
f <- function(x){
return(3*x + 2)
}
f(2)
```
### Use function() with multiple input variables
```{r}
f <- function(x, y){
x^2 + y^2
}
f(2, 2); f(2, 3); f(2, 4)
make.percent <- function(x, x.digits){
x.percent <- round(100*x, digits = x.digits)
x.show <- paste(x.percent, "%", sep = "")
return(x.show)
}
make.percent(c(0.8933, 0.2344, 0.0764, 0.7736), 0)
```
### Use function() and return multiple output
```{r}
#Introuction to list( )
X <- list(a = 3, b = c(1,2,3,4,5), c = c("A", "B"))
X
#Extract an element from a list
X <- list(a = 3, b = c(1,2,3,4,5), c = c("A", "B"))
X$a
X[[2]]
X[["c"]]
#Use function( ) and list( )
#Example: perfect number
perfect.number <- function(a, b){
k <- a
res <- c()
ans <- list()
while(k < b){
fac <- c()
for(j in 1:k){
if((k %% j) == 0) fac <- c(fac, j)
}
if(k == sum(fac[-length(fac)])) {
res <- c(res, k)
ans[[paste(k)]] <- fac[-length(fac)]
}
k <- k+1
}
ans$number <- res
return(ans)
}
#Use invisible( ) to hide some output
perfect.number <- function(a, b){
k <- a
res <- c()
ans <- list()
while(k < b){
fac <- c()
for(j in 1:k){
if((k %% j) == 0) fac <- c(fac, j)
}
if(k == sum(fac[-length(fac)])) {
res <- c(res, k)
ans[[paste(k)]] <- fac[-length(fac)]
}
k <- k+1
}
print(res)
ans$number <- res
return(invisible(ans))
}
```
# Data frame
+ Most data sets are stored in R as data frames.
+ These are like matrices, but with the columns having their own names.
+ Columns can be of different types from each other.
```{r}
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers,
more.numbers = c(4, 5, 6))
colors.and.numbers
str(colors.and.numbers)
#Types in a data frame: character
colors <- c("red", "yellow", "blue")
numbers <- c(1, 2, 3)
colors.and.numbers <- data.frame(colors, numbers,
more.numbers = c(4, 5, 6),
stringsAsFactors = FALSE)
str(colors.and.numbers)
#Add column and row vectors into a data frame
data1 <- data.frame(colors, numbers, more.numbers = c(4, 5, 6))
data1 <- rbind(data1, c("black", 3, 5))
data1
#More on data.frame(): add columns
colors <- c("red", "yellow", "blue")
number1 <- c(1, 2, 3)
number2 <- c(10, 11, 12)
colors.and.numbers <- data.frame(colors, number1) ## construct data.frame
colors.and.numbers$more.numbers <- c(4, 5, 6) ## add one column
colors.and.numbers <- cbind(colors.and.numbers, number2) ## add one column
colors.and.numbers
```
# Graphics
Given an arbitray vector x:  
1. Histogram: hist(x)  
2. Boxplot: boxplot(x)  
3. Density plot: plot(density(x))
```{r}
x.norm <- rnorm(100, mean=0, sd=1)
hist(x.norm)
boxplot(x.norm) 
plot(density(x.norm))
x <- sample(c(1:8), 100, replace=TRUE, prob=c(8:1))
x <- table(x)
names(x) <- c("A", "B", "C", "D", "E", "F", "G", "H")
barplot(x)
pie(x)
#Scatter plot is for the correlation between two variables
x <- rnorm(100)
y <- 3*x + rnorm(100, 0, 1)
plot(x, y)
X <- matrix(rnorm(30), 10, 3) ## use matrix
pairs(X, pch = 19)
X <- matrix(rnorm(50), 10, 5) ## use matrix
boxplot(X)
```

### Create a plot region with specifical items
plot(x, y,  
type = "p", # types of the plot  
pch = , # the symbol in the plot  
main = NULL, # the title of the plot  
lty = , # the style of a line  
lwd = , # the width of a line  
cex = , # the size of an text  
bty="o", # the box around the plot  
col = , # the color of an object  
xlim = NULL, ylim = NULL, # the range of an axis  
sub = NULL, # the sub-title of the plot  
xlab = NULL, ylab = NULL, # the name of an axis  
cex.main = # the size of the title  
)
```{r}
#curve(. . . )
par(mfrow = c(2, 3))
curve(exp(-x), 0, 10)
curve(dnorm(x), -10, 10)
curve(x^2, -10, 10)
curve(sin, 0, 10)
curve(3*exp(-3*x), 0, 6)
curve(5*x+2, 0, 10)
```

### Add points or lines to an existing plot
+ points(. . . ): add points
+ lines(. . . ): add lines
+ abline(. . . ): add straight lines

### Add text to an existing plot
+ text(x, y, labels = . . . )
+ legend(x, y, legend, . . . )
```{r}
x <- rnorm(50)
y <- 3*x + rnorm(50, 0, 1)
y1 <- 5*x + rnorm(50, 0, 2)
plot(x, y, type = "p", pch = 19)
abline(a = 0, b = 3)
points(x, y1, col = 4, pch = 20)
abline(a = 0, b = 5, col = 4)
text(-1.5, 2, "A", cex = 5)
text(1, -5, "B", cex = 5, col = 2)
legend("topleft", c("Y", "Y1"), col = c(1,4), lty = c(1, 1),
cex = 2)
```

### Re-arrange the region of a figure
Use par(. . . ) to creat an arbitrary region  
1. mfrow = c(2, 2): Define a multi-figure display by row  
2. mfcol = c(2, 2): Define a multi-figure display by column  
3. mai, mar = c(3, 3, 1, 1): The four margin sizes, in inches (mai), or in text line units (mar).  
Use layout(. . . )    
1. layout(matrix(c(1,2,3,3), 2, 2, byrow=TRUE))

# Rmarkdown
+ Compile a single R Markdown document to a report in different formats, such as PDF, HTML, or Word.
+ Create notebooks in which you can directly run code chunks interactively.
+ Make slides for presentations (HTML5, LaTeX Beamer, or PowerPoint).

### Some basic editings
*italics* and _italics_  
**bold** and __bold__  
superscript^2^  
~~strikethrough~~  
$A = \pi*r^{2}$  
$$B = 2 \pi r$$

### Table
Table Header  | Second Header
------------- | -------------
Table Cell    | Cell 2
Cell 3        | Cell 4

### Inline R codes
```r 2 * 3```

# Apply family
### Function: apply(. . . )
apply(X, MARGIN, FUN, . . . ) Returns a vector or array or list of values obtained by applying a function to margins of an array or a matrix.
```{r}
A <- matrix(rnorm(1500), 10, 3)
apply(A, 1, sum); rowSums(A)
apply(A, 2, sum); colSums(A)
```
### Function: lapply(. . . )
Returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.
```{r}
C <- list(a = 1:10, beta = exp(-3:3))
lapply(C, mean)
par(mfrow = c(1, 2))
lapply(C, hist)
```
### Function: sapply(. . . )
Similarlly to lapply(X, Fun, . . . ) if sapply(x, f, simplify = FALSE, USE.NAMES = FALSE)
```{r}
sapply(C, summary)
sapply(C, summary, simplify = FALSE)
```